---
title: "Replicate the Original Results using Linear Regression and Mixed-effects Model"
author: "Jie Zhou"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: show
editor_options: 
  
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(forestplot)
library(dplyr)
library(DBI)
library(DT)
library(ggplot2)
library(ggrepel)
library(mixOmics)
library(stringr)
library(VCA)
library(nlme)
library(danovir)
source("./code/data_cleaning_no_buffer.R")
source("./code/contrast.R")
options(useFancyQuotes = FALSE)
par(mar=c(1,1,1,1))
```



```{r,echo=FALSE}
library(RColorBrewer)
# colors=c("#E04C5C","#7DAF4C","#23AECE", "#FB894B", "#E7DA36",  "#187A51",
#          "#5EA4A2",  "#3D3C4E", "#4D1836", "#C51B7D",
#          "#E9A3C9",  "#B35806", "#F1A340", "#FEE08B", "#D9EF8B",
#          "#91CF60", "#C7EAE5", "#5AB4AC", "#01665E", "#E7D4E8",
#          "#AF8DC3", "#762A83","#FC0FC0","#F9C7DE","#f3a0c4")
# names(colors)=c("IgA2",      "IgA1",      "FcaR",      "FcgR2A131", "FcgR3A158", "IgG" ,      "ELISA IgG", "IgG4",      "IgG2",      "IgA"  ,     "IgG3"  ,    "IgM"    ,   "FcgR1A"  ,  "MN"    ,    "FcgR2b"  ,  "IgG1" ,
#  "HAI",       "NAI"  ,     "FcgR3b","ADCC","ADCP","ADCD")
colors = c(IgG = "#1e90ff",

                   IgA = "#3f007d",

                   IgA1 = "#C7EAE5",

                   IgA2 = "#9e9ac8",

                   IgD = "#662506",

                   IgM = "#8b7765",

                   IgG1 = "#C6E2FF",

                   IgG2 = "#CC3232",

                   IgG3 = "#4F94CD",

                   IgG4 = "#1D3700",

                   FcgR2A131 = "#157B74",

                   FcgR2b = "#CDCD00",

                   FcgR3A158 = "#722841",

                   FcgR3b = "#23AECE",

                   FcaR = "#bcbddc",
            FcRn = "#7B5D56",

                   Func="#000000",
           
           ADCC= "#80CBC4",
           ADCP="#A5D6A7",
           ADCD="#2166AC",
           MN= "#23AECE",
           HAI=  "#C7EAE5",
           NAI=  "#5AB4AC",
          "ELISA IgG"= "#7DAF4C",
           CA07="#1B4F72",
            FcgR2B = "#CDCD00",
           FcgR1A="#AA4371",

                   FcgR3A = "#722841",

                   FcgR3B = "#23AECE",
           FcgR2A = "#458B74",
          
 "ELISA"= "#7DAF4C"
           )
```





```{r,warning=FALSE,echo=FALSE}
## exclude non Lumbaa assay
device=do.call(rbind,strsplit(combined_experiments$detect_method,split="_"))
combined_experiments=cbind(combined_experiments,assay=device[,1])
 index=which(device[,1]=="LumBAA" | device[,1]=="NAI" | device[,1]=="MN" | device[,1]=="HAI" | device[,1]=="ELISA"
               | device[,1]=="ADCC" | device[,1]=="ADCP" | (device[,1]=="ADCD" & device[,3]=="dil3200" & device[,2]=="set1" ))

combined_experiments=combined_experiments[index,]
index=which(combined_experiments$detect_method=="ADCD_set1_dil3200")
combined_experiments$detect_method[index]="ADCD"



## exclude non ca data
 index=grep("CA",combined_experiments$analyte_reported)
 combined_experiments=combined_experiments[index,]
 ## exclude HA_dtm
 index=which(combined_experiments$analyte_type=="HA_dTM")
 combined_experiments=combined_experiments[-index,]
## create features
temp = paste(combined_experiments$assay, combined_experiments$analyte_reported, sep="_")
temp=ifelse(temp=="NAI_NAI_CA07","NAI_CA07", ifelse(temp=="MN_MN_CA07","MN_CA07",ifelse(temp=="HAI_HAI_CA07","HAI_CA07",
  ifelse(temp=="ADCC_ADCC_CA07","ADCC_CA07",ifelse(temp=="ADCP_ADCP_CA07","ADCP_CA07",
                                                   ifelse(temp=="ADCD_ADCD_CA07_HA","ADCD_CA07_HA",ifelse(temp=="ADCD_ADCD_CA07_HA_FL","ADCD_CA07_HA_FL",ifelse(temp=="ADCD_ADCD_CA07_HA_stalk","ADCD_CA07_HA_stalk",ifelse(temp=="ADCD_ADCD_CA07_NA","ADCD_CA07_NA",temp)))))))))
combined_experiments$temp=temp
## for each assay, assign a shape number for plotting in the following. 
assay=unique(combined_experiments$assay)
shapes=c(15,17,18)
names(shapes)=c("HASK",        "NA",         "HAFL")
```








```{r,echo=FALSE}
index=which(combined_experiments$arm_name=="Non-Pregnant 15 mcg" | combined_experiments$visit_name=="Delivery Visit" | combined_experiments$visit_name=="Visit 2")
data15=combined_experiments[-index,]
 index=which(data15$assay %in% c("LumBAA","ADCD","ELISA"))
 data15=data15[index,]
data15$arm=ifelse(data15$arm_name=="Pregnant 15 mg","P15","P30")
data15$antigen= ifelse(data15$analyte_type=="HA","HA",ifelse(data15$analyte_type=="HA_FL","HAFL", ifelse(data15$analyte_type=="HA_stalk","HASK",ifelse(data15$analyte_type=="live_virus", "LV","NA"))))
data15$value_reported=log2(data15$value_reported+1)
data15$antigen=factor(data15$antigen,levels = c("NA","HA","HAFL","HASK"))
data15$detect_reagent=factor(data15$detect_reagent,levels=c("IgG", "FcaR",       "FcgR1A", "FcgR2A131", "FcgR2b", "FcgR3A158", "FcgR3b", "IgA", "IgA1", "IgA2", "IgG1", "IgG2", "IgG3", "IgG4", "IgM"))
index=which(data15$antigen=="HA")
data15=data15[-index,]
data15$antigen=droplevels(data15$antigen)
```




```{r,echo=FALSE}

data15$sub001=ifelse(data15$subject_accession=="SUB8888067",1,0)

varPlot(form = value_reported~ arm+antigen+detect_reagent+assay, Data=data15,MeanLine=list(var=c("int", "arm", "antigen"), 
                      col=c("white", "blue", "magenta"), lwd=c(2,2,2)),
        BG=list(var="antigen", col=paste0("gray", c(70,80,90))),
        Points=list(pch=list(var="sub001", pch=c(15, 21)),
                    col=list(var="sub001", col=c("green", "black")),
                    #bg= list(var="sub001", #bg=c("yellow","white")),
                    cex=list(var="sub001", cex=c(0.6, 0.3))
                    ),
        VarLab=list(list(cex=1), list(cex=1),
                    list(cex=0.15, srt=1), list(cex=0.15)),
        YLabel=list(text="log10(value+1)(MFI)", las=0, line=3, font=2, cex=1.25),
    Title=list(main="Doseage", cex.main=1.75)
        )
data15=groupedData(value_reported~arm|subject_accession,data=data15)
data15$feature=paste0(data15$antigen,data15$detect_reagent,sep="")
contr=lmeControl(maxIter = 100,msMaxIter = 100,niterEM = 100)
```




### Mode 1 : third order model + independent error + feature-specific variance


$$infant_{ijkl}=\mu +  \alpha_j + \beta_k + \lambda_l+\gamma_{1lj}+\gamma_{2lk}+ \gamma_{3jk}+\gamma_{jkl}+ e_{ijk}, e_{ijk}\sim N(0,\sigma_{jk})$$




```{r,message=FALSE,echo=FALSE}
mainEffect2=gls(value_reported~arm+ antigen+detect_reagent+assay+ arm:antigen+arm:detect_reagent+antigen:detect_reagent+arm:antigen:detect_reagent,data = data15 ,weights = varIdent(form=~1|feature),method = "ML")
aa=summary(mainEffect2)
```

```{r,echo=FALSE,message=FALSE}
sele=list(antigen=c(2,21,22),detect_reagent=c(2,23:36),three=c(2,65:92))
names(sele$antigen)=levels(data15$antigen)
names(sele$detect_reagent)=levels(data15$detect_reagent)
coeVec=aa$coefficients
varMatrix=aa$varBeta
bb=contrast(sele=sele,coeVec = coeVec,varMatrix = varMatrix)
print("Effects of maternal immunization on features (antigen+analyte):  ")
```



```{r,echo=FALSE}
antigen=rownames(bb$effect)
antibody=colnames(bb$effect)
basic=matrix(nrow=length(antigen),ncol=length(antibody))
pp= matrix(nrow=length(antigen),ncol=length(antibody))
vv= matrix(nrow=length(antigen),ncol=length(antibody))
for (i in 1:length(antigen)) {
  for (j in 1:length(antibody)) {
    index=which(data15$antigen==antigen[i] & data15$detect_reagent==antibody[j])
    dd=data15[index,]
    x=dd$value_reported[which(dd$arm=="P15")]
        y=dd$value_reported[which(dd$arm=="P30")]
        basic[i,j]=mean(y)-mean(x)
        vv[i,j]=var(x)/length(x)+var(y)/length(y)
        pp[i,j]=2*(1-pnorm(abs(basic[i,j])/sqrt(vv[i,j])))
  }
}
```




```{r,echo=FALSE}
meanData=data.frame(antigen=rep(rownames(bb$effect),ncol(bb$effect)),antibody=rep(colnames(bb$effect),rep(nrow(bb$effect),ncol(bb$effect))) ,meffects=c(bb$effect),deffects=c(basic))


ggplot(data=meanData,aes(x=meffects,y=deffects))+
    geom_point(aes(shape=antigen,color=antibody),size=4)+
     scale_shape_manual(values = shapes)  +
    scale_color_manual(values = colors)+
   geom_abline(intercept = 0,slope = 1)+
  theme(legend.key=element_blank())+
  labs(x=bquote("model-based method"), y=bquote("original method"))+
   theme(legend.position = "right",
          legend.box = "vertical")+
   guides(color = guide_legend(nrow = 8,title="Detection Reagent"))
```


```{r,echo=FALSE}
varianceData=data.frame(antigen=rep(rownames(bb$effect),ncol(bb$effect)),antibody=rep(colnames(bb$effect),rep(nrow(bb$effect),ncol(bb$effect))) ,mvariances=c(bb$variance),dvariances=c(vv))


ggplot(data=varianceData,aes(x=mvariances,y=dvariances))+
    geom_point(aes(shape=antigen,color=antibody),size=4)+
     scale_shape_manual(values = shapes)  +
    scale_color_manual(values = colors)+
   geom_abline(intercept = 0,slope = 1)+
  theme(legend.key=element_blank())+
  labs(x=bquote("model-based method"), y=bquote("original method"))+
   theme(legend.position = "right",
          legend.box = "vertical")+
   guides(color = guide_legend(nrow = 8,title="Detection Reagent"))
```


```{r,eval=FALSE}
comparison=merge(meanData,varianceData)
comparison
```



#### model-based volcano

```{r,echo=FALSE}
volcanoData=data.frame(antigen=rep(rownames(bb$effect),ncol(bb$effect)),antibody=rep(colnames(bb$effect),rep(nrow(bb$effect),ncol(bb$effect))) ,effects=c(bb$effect),pvalue=c(bb$pvalue))
volcanoData$fdr=p.adjust(volcanoData$pvalue,method = "BH")
volcanoData=volcanoData[order(volcanoData$pvalue),]


cutoff <- ifelse(volcanoData$fdr<0.2,0.2,volcanoData$fdr)
index=max(which(volcanoData$fdr<0.2))
h=-log10(volcanoData$pvalue[index])
xx=max(abs(volcanoData$effects))

ggplot(data=volcanoData,aes(x=effects,y=-log10(pvalue)))+
    geom_point(aes(shape=antigen,color=antibody),size=4)+
     scale_shape_manual(values = shapes)  +
    scale_color_manual(values = colors)+
   theme_classic()+
  theme(legend.key=element_blank())+
  geom_hline(yintercept=h,linetype="dashed")+
  geom_vline(xintercept=0,linetype="dashed")+
  labs(x=bquote("Fold Change"), y=bquote(-Log[10](p-value)))+
  #ggtitle( paste0("Day0: NP vs P15(R^2=",round(rsqr,2),")" ) )+
   theme(legend.position = "right",
          legend.box = "vertical")+
   guides(color = guide_legend(nrow = 8,title="Detection Reagent"))+
    geom_text(aes(xx/2,h,label = paste("FDR=", 0.2, sep=""), vjust = -1),show.legend = FALSE)
```


#### original volcano

```{r,echo=FALSE}
volcanoData=data.frame(antigen=rep(rownames(bb$effect),ncol(bb$effect)),antibody=rep(colnames(bb$effect),rep(nrow(bb$effect),ncol(bb$effect))) ,effects=c(basic),pvalue=c(pp))
volcanoData$fdr=p.adjust(volcanoData$pvalue,method = "BH")
volcanoData=volcanoData[order(volcanoData$pvalue),]


cutoff <- ifelse(volcanoData$fdr<0.2,0.2,volcanoData$fdr)
index=max(which(volcanoData$fdr<0.2))
h=-log10(volcanoData$pvalue[index])
xx=max(abs(volcanoData$effects))

ggplot(data=volcanoData,aes(x=effects,y=-log10(pvalue)))+
    geom_point(aes(shape=antigen,color=antibody),size=4)+
     scale_shape_manual(values = shapes)  +
    scale_color_manual(values = colors)+
   theme_classic()+
  #xlim(-1,1)+
  theme(legend.key=element_blank())+
  geom_hline(yintercept=h,linetype="dashed")+
  geom_vline(xintercept=0,linetype="dashed")+
  labs(x=bquote("Fold Change"), y=bquote(-Log[10](p-value)))+
    #ggtitle( paste0("Day0: P15 vs P30(R^2=",round(rsqr,2),")" ) )+
   theme(legend.position = "right",
          legend.box = "vertical")+
   guides(color = guide_legend(nrow = 8,title="Detection Reagent"))+
    geom_text(aes(xx/2,h,label = paste("FDR=", 0.2, sep=""), vjust = -1),show.legend = FALSE)
```




### Model 2: third order model + random intercept + feature-specific variance


$$infant_{ijkl}=\mu +  \alpha_j + \beta_k + \lambda_l+\gamma_{1lj}+\gamma_{2lk}+ \gamma_{3jk}+\gamma_{4jkl}+u_i+ e_{ijk}, e_{ijk}\sim N(0,\sigma_{jk})$$

```{r,message=FALSE,echo=FALSE}
mainEffect3=lme(value_reported~arm+ antigen+detect_reagent+assay+ arm:antigen+arm:detect_reagent+antigen:detect_reagent+arm:antigen:detect_reagent,random=pdDiag(~1),weights = varIdent(form = ~1|feature),data = data15,method = "ML",control = contr)
aa=summary(mainEffect3)
```



```{r,echo=FALSE,message=FALSE}
sele=list(antigen=c(2,21,22),detect_reagent=c(2,23:36),three=c(2,65:92))
names(sele$antigen)=levels(data15$antigen)
names(sele$detect_reagent)=levels(data15$detect_reagent)
coeVec=aa$coefficients$fixed
varMatrix=aa$varFix
bb=contrast(sele=sele,coeVec = coeVec,varMatrix = varMatrix)
print("Effects of maternal immunization on features (antigen+analyte):  ")
```




```{r,echo=FALSE}
meanData=data.frame(antigen=rep(rownames(bb$effect),ncol(bb$effect)),antibody=rep(colnames(bb$effect),rep(nrow(bb$effect),ncol(bb$effect))) ,meffects=c(bb$effect),deffects=c(basic))


ggplot(data=meanData,aes(x=meffects,y=deffects))+
    geom_point(aes(shape=antigen,color=antibody),size=4)+
     scale_shape_manual(values = shapes)  +
    scale_color_manual(values = colors)+
   geom_abline(intercept = 0,slope = 1)+
  theme(legend.key=element_blank())+
  labs(x=bquote("model-based method"), y=bquote("original method"))+
   theme(legend.position = "right",
          legend.box = "vertical")+
   guides(color = guide_legend(nrow = 8,title="Detection Reagent"))
```


```{r,echo=FALSE}
varianceData=data.frame(antigen=rep(rownames(bb$effect),ncol(bb$effect)),antibody=rep(colnames(bb$effect),rep(nrow(bb$effect),ncol(bb$effect))) ,mvariances=c(bb$variance),dvariances=c(vv))


ggplot(data=varianceData,aes(x=mvariances,y=dvariances))+
    geom_point(aes(shape=antigen,color=antibody),size=4)+
     scale_shape_manual(values = shapes)  +
    scale_color_manual(values = colors)+
   geom_abline(intercept = 0,slope = 1)+
  theme(legend.key=element_blank())+
  labs(x=bquote("model-based method"), y=bquote("original method"))+
   theme(legend.position = "right",
          legend.box = "vertical")+
   guides(color = guide_legend(nrow = 8,title="Detection Reagent"))
```


```{r,eval=FALSE}
comparison=merge(meanData,varianceData)
comparison
```





#### model-based volcano

```{r,echo=FALSE}
volcanoData=data.frame(antigen=rep(rownames(bb$effect),ncol(bb$effect)),antibody=rep(colnames(bb$effect),rep(nrow(bb$effect),ncol(bb$effect))) ,effects=c(bb$effect),pvalue=c(bb$pvalue))
volcanoData$fdr=p.adjust(volcanoData$pvalue,method = "BH")
volcanoData=volcanoData[order(volcanoData$pvalue),]


cutoff <- ifelse(volcanoData$fdr<0.2,0.2,volcanoData$fdr)
index=max(which(volcanoData$fdr<0.2))
h=-log10(volcanoData$pvalue[index])
xx=max(abs(volcanoData$effects))

ggplot(data=volcanoData,aes(x=effects,y=-log10(pvalue)))+
    geom_point(aes(shape=antigen,color=antibody),size=4)+
     scale_shape_manual(values = shapes)  +
    scale_color_manual(values = colors)+
   theme_classic()+
  theme(legend.key=element_blank())+
  geom_hline(yintercept=h,linetype="dashed")+
  geom_vline(xintercept=0,linetype="dashed")+
  labs(x=bquote("Fold Change"), y=bquote(-Log[10](p-value)))+
  #ggtitle( paste0("Day0: NP vs P15(R^2=",round(rsqr,2),")" ) )+
   theme(legend.position = "right",
          legend.box = "vertical")+
   guides(color = guide_legend(nrow = 8,title="Detection Reagent"))+
    geom_text(aes(xx/2,h,label = paste("FDR=", 0.2, sep=""), vjust = -1),show.legend = FALSE)
```


#### original volcano

```{r,echo=FALSE}
volcanoData=data.frame(antigen=rep(rownames(bb$effect),ncol(bb$effect)),antibody=rep(colnames(bb$effect),rep(nrow(bb$effect),ncol(bb$effect))) ,effects=c(basic),pvalue=c(pp))
volcanoData$fdr=p.adjust(volcanoData$pvalue,method = "BH")
volcanoData=volcanoData[order(volcanoData$pvalue),]


cutoff <- ifelse(volcanoData$fdr<0.2,0.2,volcanoData$fdr)
index=max(which(volcanoData$fdr<0.2))
h=-log10(volcanoData$pvalue[index])
xx=max(abs(volcanoData$effects))

ggplot(data=volcanoData,aes(x=effects,y=-log10(pvalue)))+
    geom_point(aes(shape=antigen,color=antibody),size=4)+
     scale_shape_manual(values = shapes)  +
    scale_color_manual(values = colors)+
   theme_classic()+
  #xlim(-1,1)+
  theme(legend.key=element_blank())+
  geom_hline(yintercept=h,linetype="dashed")+
  geom_vline(xintercept=0,linetype="dashed")+
  labs(x=bquote("Fold Change"), y=bquote(-Log[10](p-value)))+
    #ggtitle( paste0("Day0: P15 vs P30(R^2=",round(rsqr,2),")" ) )+
   theme(legend.position = "right",
          legend.box = "vertical")+
   guides(color = guide_legend(nrow = 8,title="Detection Reagent"))+
    geom_text(aes(xx/2,h,label = paste("FDR=", 0.2, sep=""), vjust = -1),show.legend = FALSE)
```

#### forest plot


```{r,echo=FALSE}
ss=as.data.frame(aa$tTable[c(sele[[1]],sele[[2]][-1],sele[[3]][-1]),c(1,2)])
ss$lower=ss[,1]-1.96*ss[,2]
ss$upper=ss[,1]+1.96*ss[,2]
ss$name=rownames(ss)
rownames(ss)=c()
```

```{r,echo=FALSE}
base_data=tibble::tibble(mean=ss$Value,lower=ss$lower,upper=ss$upper,interaction=substr(ss$name,14,nchar(ss$name)))
base_data$interaction=c("ref:na+igg","HAFL","HASK",substr(ss$name[4:17],22,nchar(ss$name[4:17])),paste(substr(ss$name[18:45],15,18),substr(ss$name[18:45],34,50),sep =":"))
```




```{r,echo=FALSE}
ss2=as.data.frame(aa$tTable[c(3:18,37:64),c(1,2)])
ss2$lower=ss2[,1]-1.96*ss2[,2]
ss2$upper=ss2[,1]+1.96*ss2[,2]
ss2$name=rownames(ss2)
rownames(ss2)=c()
```

```{r,echo=FALSE}
base_data2=tibble::tibble(mean=ss2$Value,lower=ss2$lower,upper=ss2$upper,interaction=substr(ss2$name,14,nchar(ss2$name)))
base_data2$interaction=c("HAFL","HASK",substr(ss2$name[4:17],nchar(ss2$name[4:17])-3,nchar(ss2$name[4:17])),paste(substr(ss2$name[18:45],8,11),substr(ss2$name[18:45],27,50),sep =":"))
```



##### full effects decomposition

```{r,echo=FALSE}
base_data3=rbind(base_data,base_data2)
```

```{r,echo=FALSE}
base_data3 |>
  forestplot(labeltext = c(interaction),txt_gp=fpTxtGp(cex=0.5,label = gpar(fontfamily = "HersheySerif")),
             clip = c(-6, 6)) |>
  fp_set_style(box = "royalblue",
               line = "darkblue",
               summary = "royalblue") |> 
  fp_add_header(interaction= c("Direct effects")) |>
  # fp_append_row(mean  = 0.531,
  #               lower = 0.386,
  #               upper = 0.731,
  #               study = "Summary",
  #               OR = "0.53",
  #               is.summary = TRUE) |>
  fp_set_zebra_style("#EFEFEF")
```








### Model 3: third order model + random intercept/antigen + homoscedastic error


$$infant_{ijkl}=\mu +  \alpha_j + \beta_k + \lambda_l+\gamma_{1lj}+\gamma_{2lk}+ \gamma_{3jk}+\gamma_{4jkl}+u_i+ u_{ij}+u_{ik}+ e_{ijk}, e_{ijk}\sim N(0,\sigma^2)$$

```{r,message=FALSE,echo=FALSE}
mainEffect3=lme(value_reported~arm+ antigen+detect_reagent+assay+ arm:antigen+arm:detect_reagent+antigen:detect_reagent+arm:antigen:detect_reagent,random=pdDiag(~antigen),weights = varIdent(form = ~1),data = data15,method = "ML",control = contr)
aa=summary(mainEffect3)
```



```{r,echo=FALSE,message=FALSE}
sele=list(antigen=c(2,21,22),detect_reagent=c(2,23:36),three=c(2,65:92))
names(sele$antigen)=levels(data15$antigen)
names(sele$detect_reagent)=levels(data15$detect_reagent)
coeVec=aa$coefficients$fixed
varMatrix=aa$varFix
bb=contrast(sele=sele,coeVec = coeVec,varMatrix = varMatrix)
print("Effects of maternal immunization on features (antigen+analyte):  ")
```




```{r,echo=FALSE}
meanData=data.frame(antigen=rep(rownames(bb$effect),ncol(bb$effect)),antibody=rep(colnames(bb$effect),rep(nrow(bb$effect),ncol(bb$effect))) ,meffects=c(bb$effect),deffects=c(basic))


ggplot(data=meanData,aes(x=meffects,y=deffects))+
    geom_point(aes(shape=antigen,color=antibody),size=4)+
     scale_shape_manual(values = shapes)  +
    scale_color_manual(values = colors)+
   geom_abline(intercept = 0,slope = 1)+
  theme(legend.key=element_blank())+
  labs(x=bquote("model-based method"), y=bquote("original method"))+
   theme(legend.position = "right",
          legend.box = "vertical")+
   guides(color = guide_legend(nrow = 8,title="Detection Reagent"))
```


```{r,echo=FALSE}
varianceData=data.frame(antigen=rep(rownames(bb$effect),ncol(bb$effect)),antibody=rep(colnames(bb$effect),rep(nrow(bb$effect),ncol(bb$effect))) ,mvariances=c(bb$variance),dvariances=c(vv))


ggplot(data=varianceData,aes(x=mvariances,y=dvariances))+
    geom_point(aes(shape=antigen,color=antibody),size=4)+
     scale_shape_manual(values = shapes)  +
    scale_color_manual(values = colors)+
   geom_abline(intercept = 0,slope = 1)+
  theme(legend.key=element_blank())+
  labs(x=bquote("model-based method"), y=bquote("original method"))+
   theme(legend.position = "right",
          legend.box = "vertical")+
   guides(color = guide_legend(nrow = 8,title="Detection Reagent"))
```


```{r,eval=FALSE}
comparison=merge(meanData,varianceData)
comparison
```





#### model-based volcano

```{r,echo=FALSE}
volcanoData=data.frame(antigen=rep(rownames(bb$effect),ncol(bb$effect)),antibody=rep(colnames(bb$effect),rep(nrow(bb$effect),ncol(bb$effect))) ,effects=c(bb$effect),pvalue=c(bb$pvalue))
volcanoData$fdr=p.adjust(volcanoData$pvalue,method = "BH")
volcanoData=volcanoData[order(volcanoData$pvalue),]


cutoff <- ifelse(volcanoData$fdr<0.2,0.2,volcanoData$fdr)
index=max(which(volcanoData$fdr<0.2))
h=-log10(volcanoData$pvalue[index])
xx=max(abs(volcanoData$effects))

ggplot(data=volcanoData,aes(x=effects,y=-log10(pvalue)))+
    geom_point(aes(shape=antigen,color=antibody),size=4)+
     scale_shape_manual(values = shapes)  +
    scale_color_manual(values = colors)+
   theme_classic()+
  theme(legend.key=element_blank())+
  geom_hline(yintercept=h,linetype="dashed")+
  geom_vline(xintercept=0,linetype="dashed")+
  labs(x=bquote("Fold Change"), y=bquote(-Log[10](p-value)))+
  #ggtitle( paste0("Day0: NP vs P15(R^2=",round(rsqr,2),")" ) )+
   theme(legend.position = "right",
          legend.box = "vertical")+
   guides(color = guide_legend(nrow = 8,title="Detection Reagent"))+
    geom_text(aes(xx/2,h,label = paste("FDR=", 0.2, sep=""), vjust = -1),show.legend = FALSE)
```


#### original volcano

```{r,echo=FALSE}
volcanoData=data.frame(antigen=rep(rownames(bb$effect),ncol(bb$effect)),antibody=rep(colnames(bb$effect),rep(nrow(bb$effect),ncol(bb$effect))) ,effects=c(basic),pvalue=c(pp))
volcanoData$fdr=p.adjust(volcanoData$pvalue,method = "BH")
volcanoData=volcanoData[order(volcanoData$pvalue),]


cutoff <- ifelse(volcanoData$fdr<0.2,0.2,volcanoData$fdr)
index=max(which(volcanoData$fdr<0.2))
h=-log10(volcanoData$pvalue[index])
xx=max(abs(volcanoData$effects))

ggplot(data=volcanoData,aes(x=effects,y=-log10(pvalue)))+
    geom_point(aes(shape=antigen,color=antibody),size=4)+
     scale_shape_manual(values = shapes)  +
    scale_color_manual(values = colors)+
   theme_classic()+
  #xlim(-1,1)+
  theme(legend.key=element_blank())+
  geom_hline(yintercept=h,linetype="dashed")+
  geom_vline(xintercept=0,linetype="dashed")+
  labs(x=bquote("Fold Change"), y=bquote(-Log[10](p-value)))+
    #ggtitle( paste0("Day0: P15 vs P30(R^2=",round(rsqr,2),")" ) )+
   theme(legend.position = "right",
          legend.box = "vertical")+
   guides(color = guide_legend(nrow = 8,title="Detection Reagent"))+
    geom_text(aes(xx/2,h,label = paste("FDR=", 0.2, sep=""), vjust = -1),show.legend = FALSE)
```




### Model 4: third order model + random intercept/antibody + homoscedastic error


$$infant_{ijkl}=\mu +  \alpha_j + \beta_k + \lambda_l+\gamma_{1lj}+\gamma_{2lk}+ \gamma_{3jk}+\gamma_{4jkl}+u_i+ u_{ij}+u_{ik}+ e_{ijk}, e_{ijk}\sim N(0,\sigma^2)$$

```{r,message=FALSE,echo=FALSE}
mainEffect3=lme(value_reported~arm+ antigen+detect_reagent+assay+ arm:antigen+arm:detect_reagent+antigen:detect_reagent+arm:antigen:detect_reagent,random=pdDiag(~detect_reagent),weights = varIdent(form = ~1),data = data15,method = "ML",control = contr)
aa=summary(mainEffect3)
```



```{r,echo=FALSE,message=FALSE}
sele=list(antigen=c(2,21,22),detect_reagent=c(2,23:36),three=c(2,65:92))
names(sele$antigen)=levels(data15$antigen)
names(sele$detect_reagent)=levels(data15$detect_reagent)
coeVec=aa$coefficients$fixed
varMatrix=aa$varFix
bb=contrast(sele=sele,coeVec = coeVec,varMatrix = varMatrix)
print("Effects of maternal immunization on features (antigen+analyte):  ")
```




```{r,echo=FALSE}
meanData=data.frame(antigen=rep(rownames(bb$effect),ncol(bb$effect)),antibody=rep(colnames(bb$effect),rep(nrow(bb$effect),ncol(bb$effect))) ,meffects=c(bb$effect),deffects=c(basic))


ggplot(data=meanData,aes(x=meffects,y=deffects))+
    geom_point(aes(shape=antigen,color=antibody),size=4)+
     scale_shape_manual(values = shapes)  +
    scale_color_manual(values = colors)+
   geom_abline(intercept = 0,slope = 1)+
  theme(legend.key=element_blank())+
  labs(x=bquote("model-based method"), y=bquote("original method"))+
   theme(legend.position = "right",
          legend.box = "vertical")+
   guides(color = guide_legend(nrow = 8,title="Detection Reagent"))
```


```{r,echo=FALSE}
varianceData=data.frame(antigen=rep(rownames(bb$effect),ncol(bb$effect)),antibody=rep(colnames(bb$effect),rep(nrow(bb$effect),ncol(bb$effect))) ,mvariances=c(bb$variance),dvariances=c(vv))


ggplot(data=varianceData,aes(x=mvariances,y=dvariances))+
    geom_point(aes(shape=antigen,color=antibody),size=4)+
     scale_shape_manual(values = shapes)  +
    scale_color_manual(values = colors)+
   geom_abline(intercept = 0,slope = 1)+
  theme(legend.key=element_blank())+
  labs(x=bquote("model-based method"), y=bquote("original method"))+
   theme(legend.position = "right",
          legend.box = "vertical")+
   guides(color = guide_legend(nrow = 8,title="Detection Reagent"))
```


```{r,eval=FALSE}
comparison=merge(meanData,varianceData)
comparison
```





#### model-based volcano

```{r,echo=FALSE}
volcanoData=data.frame(antigen=rep(rownames(bb$effect),ncol(bb$effect)),antibody=rep(colnames(bb$effect),rep(nrow(bb$effect),ncol(bb$effect))) ,effects=c(bb$effect),pvalue=c(bb$pvalue))
volcanoData$fdr=p.adjust(volcanoData$pvalue,method = "BH")
volcanoData=volcanoData[order(volcanoData$pvalue),]


cutoff <- ifelse(volcanoData$fdr<0.2,0.2,volcanoData$fdr)
index=max(which(volcanoData$fdr<0.2))
h=-log10(volcanoData$pvalue[index])
xx=max(abs(volcanoData$effects))

ggplot(data=volcanoData,aes(x=effects,y=-log10(pvalue)))+
    geom_point(aes(shape=antigen,color=antibody),size=4)+
     scale_shape_manual(values = shapes)  +
    scale_color_manual(values = colors)+
   theme_classic()+
  theme(legend.key=element_blank())+
  geom_hline(yintercept=h,linetype="dashed")+
  geom_vline(xintercept=0,linetype="dashed")+
  labs(x=bquote("Fold Change"), y=bquote(-Log[10](p-value)))+
  #ggtitle( paste0("Day0: NP vs P15(R^2=",round(rsqr,2),")" ) )+
   theme(legend.position = "right",
          legend.box = "vertical")+
   guides(color = guide_legend(nrow = 8,title="Detection Reagent"))+
    geom_text(aes(xx/2,h,label = paste("FDR=", 0.2, sep=""), vjust = -1),show.legend = FALSE)
```


#### original volcano

```{r,echo=FALSE}
volcanoData=data.frame(antigen=rep(rownames(bb$effect),ncol(bb$effect)),antibody=rep(colnames(bb$effect),rep(nrow(bb$effect),ncol(bb$effect))) ,effects=c(basic),pvalue=c(pp))
volcanoData$fdr=p.adjust(volcanoData$pvalue,method = "BH")
volcanoData=volcanoData[order(volcanoData$pvalue),]


cutoff <- ifelse(volcanoData$fdr<0.2,0.2,volcanoData$fdr)
index=max(which(volcanoData$fdr<0.2))
h=-log10(volcanoData$pvalue[index])
xx=max(abs(volcanoData$effects))

ggplot(data=volcanoData,aes(x=effects,y=-log10(pvalue)))+
    geom_point(aes(shape=antigen,color=antibody),size=4)+
     scale_shape_manual(values = shapes)  +
    scale_color_manual(values = colors)+
   theme_classic()+
  #xlim(-1,1)+
  theme(legend.key=element_blank())+
  geom_hline(yintercept=h,linetype="dashed")+
  geom_vline(xintercept=0,linetype="dashed")+
  labs(x=bquote("Fold Change"), y=bquote(-Log[10](p-value)))+
    #ggtitle( paste0("Day0: P15 vs P30(R^2=",round(rsqr,2),")" ) )+
   theme(legend.position = "right",
          legend.box = "vertical")+
   guides(color = guide_legend(nrow = 8,title="Detection Reagent"))+
    geom_text(aes(xx/2,h,label = paste("FDR=", 0.2, sep=""), vjust = -1),show.legend = FALSE)
```

#### forest plot


```{r,echo=FALSE}
ss=as.data.frame(aa$tTable[c(sele[[1]],sele[[2]][-1],sele[[3]][-1]),c(1,2)])
ss$lower=ss[,1]-1.96*ss[,2]
ss$upper=ss[,1]+1.96*ss[,2]
ss$name=rownames(ss)
rownames(ss)=c()
```

```{r,echo=FALSE}
base_data=tibble::tibble(mean=ss$Value,lower=ss$lower,upper=ss$upper,interaction=substr(ss$name,14,nchar(ss$name)))
base_data$interaction=c("ref:na+igg","HAFL","HASK",substr(ss$name[4:17],22,nchar(ss$name[4:17])),paste(substr(ss$name[18:45],15,18),substr(ss$name[18:45],34,50),sep =":"))
```




```{r,echo=FALSE}
ss2=as.data.frame(aa$tTable[c(3:18,37:64),c(1,2)])
ss2$lower=ss2[,1]-1.96*ss2[,2]
ss2$upper=ss2[,1]+1.96*ss2[,2]
ss2$name=rownames(ss2)
rownames(ss2)=c()
```

```{r,echo=FALSE}
base_data2=tibble::tibble(mean=ss2$Value,lower=ss2$lower,upper=ss2$upper,interaction=substr(ss2$name,14,nchar(ss2$name)))
base_data2$interaction=c("HAFL","HASK",substr(ss2$name[4:17],nchar(ss2$name[4:17])-3,nchar(ss2$name[4:17])),paste(substr(ss2$name[18:45],8,11),substr(ss2$name[18:45],27,50),sep =":"))
```



##### full effects decomposition

```{r,echo=FALSE}
base_data3=rbind(base_data,base_data2)
```

```{r,echo=FALSE}
base_data3 |>
  forestplot(labeltext = c(interaction),txt_gp=fpTxtGp(cex=0.5,label = gpar(fontfamily = "HersheySerif")),
             clip = c(-6, 6)) |>
  fp_set_style(box = "royalblue",
               line = "darkblue",
               summary = "royalblue") |> 
  fp_add_header(interaction= c("Direct effects")) |>
  # fp_append_row(mean  = 0.531,
  #               lower = 0.386,
  #               upper = 0.731,
  #               study = "Summary",
  #               OR = "0.53",
  #               is.summary = TRUE) |>
  fp_set_zebra_style("#EFEFEF")
```








